---
title: "Deprivation"
author: "Elliot Meador"
date: "22/10/2019"
output: word_document
editor_options: 
  chunk_output_type: console
---

```{r setup, echo=FALSE, warning=F, message=F}
knitr::opts_chunk$set(echo = TRUE)
```

```{r data-library, echo=F, message=F, warning=F, results='hide'}
library(ggplot2)
library(tidyverse)
library(janitor)
library(openxlsx)
library(glue)
library(flextable)
library(scales)
library(ggrepel)

library(sf)
library(tmap)
library(tmaptools)


source('/Users/emeador/OneDrive - SRUC/all_functions.R')
    ## Deprivation Data

    dep_england <- st_read('/Users/emeador/OneDrive - SRUC/Data/deprivation/england/english_deprivation_score_sf.shp')
    
    dep_scotland <- st_read('/Users/emeador/OneDrive - SRUC/Data/deprivation/scotland/scottish_deprivation_score_sf.shp')
    
scotland_dep_df <- read_csv('/Users/emeador/OneDrive - SRUC/Data/deprivation/scotland/simd2016_withgeog/simd2016_withinds.csv') %>% 
  clean_names()

england_dep_df <- read.xlsx('/Users/emeador/OneDrive - SRUC/Data/deprivation/england/england_dep_scores_2019.xlsx', sheet = 'IoD2019 Scores') %>% 
  as_tibble() %>% 
  clean_names()


   


```


```{r plot-defaults, echo=F, message=F, warning=F}

pop_theme <- theme_bw()+
    theme(text = element_text(size = 13.5, 
                              color = 'black'), 
          panel.grid.minor = element_blank(), 
          axis.title.y = element_text(angle = 0, hjust = 1, face = 'italic'),
          plot.margin = margin(1,1,1,1, 'cm'), 
          strip.background = element_blank(), 
          strip.text.y = element_text(angle = 0, hjust = 0), 
          legend.position = 'bottom', 
          plot.caption = element_text(hjust = 0, face= "italic"), #Default is hjust=1
        plot.title.position = "plot", #NEW parameter. Apply for subtitle too.
        plot.caption.position =  "plot")



case_study_cols <- c(Spectral[c(1, 7, 11)], 
                     adjustcolor(Greys[2], .25))

names(case_study_cols) <- c("Na h-Eileanan Siar", 
                            "Northumberland",
                            "Perth and Kinross", 
                            "Other")
```

# Data
Data on deprivation comes from the English Indix of Multiple Deprivation (https://www.gov.uk/government/statistics/english-indices-of-deprivation-2019) and the Scottish Index of Multiple Deprivation (https://simd.scot/2016/#/simd2016/BTTTFTT/9/-4.0000/55.9000/). These data are aggregated at a similar small-area output (SMO), which allows us to get more granularity from the data across England and Scotland. 

The scores, which are used in calculating the index, are used -- ***not the index ranking***. For instance, one score used to determine income deprivation is the percentatge of people within a SMO that are income deprived. Index rankings themselves are not appropriate for comparing places across the two indexes as they are relational within countries (not between). 


## Employment Deprivation 
```{r employment-overview-scratch, echo=F, warning=F, message=F}




dep_england_tidy <- england_dep_df %>% 
  select(name = local_authority_district_name_2019, 
         income_rate = income_score_rate, 
         employment_rate = employment_score_rate) %>% 
  mutate(case_study = case_when(
    str_detect(name, 'Northumberland')~'Northumberland', 
    T ~ 'Rest of England-Scotland'
  )) 



dep_scotland_tidy <- scotland_dep_df %>% 
  select(name = council_area, 
         employment_rate, 
         income_rate) %>% 
  mutate_if(is.factor, as.character) %>% 
  mutate(employment_rate = parse_number(employment_rate)/100, 
         income_rate = parse_number(income_rate)/100) %>% 
  mutate(case_study = case_when(
    str_detect(name, 'Perth_and_Kinross')~'Perth and Kinross', 
    str_detect(name, 'Na_h-Eileanan_an_Iar')~'Na h-Eileanan Siar',
    T ~ 'Rest of England-Scotland'
  
  )) 


income_employ_dep <- dep_england_tidy %>% 
  bind_rows(dep_scotland_tidy)

```

```{r employment-map-scratch, echo=FALSE, warning=F, message=F}



northumberland_sf <-  dep_england %>% 
    filter(str_detect(lso11nm, 'Northumberland'))


scotland_sf <- dep_scotland %>% 
    select_if(is.factor) %>% 
    mutate_at(vars(incm_rt:ncntrlht_r), 
              list(~parse_number(as.character(.))/100))


perthshire <- scotland_sf %>% 
    filter(str_detect(concl_r, 
                      'Perth_and_Kinross'))  %>% 
    select(Name, contains('emp'), 
           income_rate = incm_rt, 
           over_crowd_rate = ovrcrwdd_r) %>% 
    rename(employment_rate = emplymnt_r)


Na_h_Eileanan_an_Iar<- scotland_sf %>% 
    filter(str_detect(concl_r, 
                      'Na_h-Eileanan_an_Iar'))  %>% 
    select(Name, contains('emp'), 
           income_rate = incm_rt, 
           over_crowd_rate = ovrcrwdd_r) %>% 
    rename(employment_rate = emplymnt_r)



northumberland_employment <- northumberland_sf %>% 
    select(Name = lso11nm, contains('emp'), 
           income_rate = incm_dprvt) %>% 
    rename(employment_rate = emp__)



place_names <- c('Perth and Kinross', 
       'Outer Hebrides', 
       'Northumberland')


```

### Employment deprivation definitions


```{r employment-table, echo=F, warning=F, message=F}

income_employ_dep %>% 
  group_by(case_study) %>% 
  summarise(Median = median(employment_rate, na.rm = T), 
          Mean = mean(employment_rate, na.rm = T), 
          Minimum = min(employment_rate, na.rm = T), 
          Maximum = max(employment_rate, na.rm = T)) %>% 
  mutate_if(is.numeric, list(~percent(.))) %>% 
  rename('Case study area' = case_study) %>% 
  flextable(cwidth = c(2,1,1,1,1)) %>% 
  set_caption('Proportion of Employment Deprivation in case study ares')
```


```{r employment-map, echo=FALSE, warning=F, message=F, fig.width=16, fig.height=6}

employment_ls <- list(perthshire,
                      Na_h_Eileanan_an_Iar,
                      northumberland_employment)

tmap_style("gray")
title_ls <- glue('{place_names}')


employment_maps <- map(1:3, function(x){
  tm_shape(employment_ls[[x]]) +
    tm_fill('employment_rate',
            title  = 'Unemployment Rate',
            palette = '-YlGnBu',
            style = 'log10',
            legend.format=
                list(fun=function(x) {paste0(round(x, 2)*100, "%")})) +
    tm_scale_bar()+
    tm_layout(title = title_ls[[x]],
              legend.frame = T,
              legend.bg.color = 'white',
              asp = 1)
})

tmap_arrange(employment_maps)


employment_map_path <- glue('/Users/emeador/OneDrive - SRUC/Twenty_five_years_on/png/maps/employment/{place_names}_income.png')

map2(employment_maps, employment_map_path, function(x, y){
  tmap_save(x,
            filename = y, 
              width = 4.5, 
            height = 4.5)
}) 

``` 


## Income Deprivation 

```{r income-table, echo=F, warning=F, message=F}
dep_england_tidy %>% 
  bind_rows(dep_scotland_tidy) %>% 
  group_by(case_study) %>% 
  summarise(Median = median(income_rate, na.rm = T), 
            Mean = mean(income_rate, na.rm = T), 
            Minimum = min(income_rate, na.rm = T), 
            Maximum = max(income_rate, na.rm = T)) %>% 
  mutate_if(is.numeric, list(~percent(.))) %>% 
  rename('Case study area' = case_study) %>% 
  flextable(cwidth = c(2,1,1,1,1)) %>% 
  set_caption('Proportion of Employment Deprivation in case study ares')


```


```{r income-map, echo=FALSE, warning=F, message=F, fig.width=16, fig.height=6}

title_ls <- glue('{place_names}')



employment_ls <- list(perthshire,
                      Na_h_Eileanan_an_Iar,
                      northumberland_employment)
tmap_style("gray")

income_maps <- map(1:3, function(x){
  tm_shape(employment_ls[[x]]) +
    tm_fill('income_rate',
            title  = 'Income Deprivation\nRate',
            palette = '-YlGnBu',
            style = 'log10',
            legend.format=
                list(fun=function(x) {paste0(round(x, 2)*100, "%")})) +
    tm_scale_bar()+
    tm_layout(title = title_ls[[x]],
              legend.frame = T,
              legend.bg.color = 'white',
              asp = 1)
})

tmap_arrange(income_maps)

 income_map_path <- glue('/Users/emeador/OneDrive - SRUC/Twenty_five_years_on/png/maps/income/{place_names}_income.png')

 map2(income_maps, income_map_path, function(x, y){
   tmap_save(x,
             filename = y,
             width = 4.5,
             height = 4.5)
 })

```











