---
title: "Deprivation"
author: "Elliot Meador"
date: "22/10/2019"
output: word_document
editor_options: 
  chunk_output_type: console
---

```{r setup, echo=FALSE, warning=F, message=F}
knitr::opts_chunk$set(echo = TRUE)
```

```{r data-library, echo=F, message=F, warning=F, results='hide'}
library(ggplot2)
library(tidyverse)
library(janitor)
library(openxlsx)
library(glue)
library(flextable)
library(scales)
library(ggrepel)

library(sf)
library(tmap)
library(tmaptools)

<<<<<<< HEAD
    ## Deprivation Data

    dep_england <- st_read('~/OneDrive - SRUC/Data/deprivation/england/english_deprivation_score_sf.shp')
    
    dep_scotland <- st_read('~/OneDrive - SRUC/Data/deprivation/scotland/scottish_deprivation_score_sf.shp')
    
scotland_dep_df <- read_csv('~/OneDrive - SRUC/Data/deprivation/scotland/simd2016_withgeog/simd2016_withinds.csv') %>% 
  clean_names()

england_dep_df <- read.xlsx('~/OneDrive - SRUC/Data/deprivation/england/england_dep_scores_2019.xlsx', sheet = 'IoD2019 Scores') %>% 
=======

source('/Users/emeador/OneDrive - SRUC/all_functions.R')
    ## Deprivation Data

    dep_england <- st_read('/Users/emeador/OneDrive - SRUC/Data/deprivation/england/english_deprivation_score_sf.shp')
    
    dep_scotland <- st_read('/Users/emeador/OneDrive - SRUC/Data/deprivation/scotland/scottish_deprivation_score_sf.shp')
    
scotland_dep_df <- read_csv('/Users/emeador/OneDrive - SRUC/Data/deprivation/scotland/simd2016_withgeog/simd2016_withinds.csv') %>% 
  clean_names()

england_dep_df <- read.xlsx('/Users/emeador/OneDrive - SRUC/Data/deprivation/england/england_dep_scores_2019.xlsx', sheet = 'IoD2019 Scores') %>% 
>>>>>>> 15fdc92a8343fff8c1805198b3b61445fcc1591e
  as_tibble() %>% 
  clean_names()


<<<<<<< HEAD
source('~/OneDrive - SRUC/all_functions.R')

    
    


=======
   
>>>>>>> 15fdc92a8343fff8c1805198b3b61445fcc1591e


```


```{r plot-defaults, echo=F, message=F, warning=F}

pop_theme <- theme_bw()+
    theme(text = element_text(size = 13.5, 
                              color = 'black'), 
          panel.grid.minor = element_blank(), 
          axis.title.y = element_text(angle = 0, hjust = 1, face = 'italic'),
          plot.margin = margin(1,1,1,1, 'cm'), 
          strip.background = element_blank(), 
          strip.text.y = element_text(angle = 0, hjust = 0), 
          legend.position = 'bottom', 
          plot.caption = element_text(hjust = 0, face= "italic"), #Default is hjust=1
        plot.title.position = "plot", #NEW parameter. Apply for subtitle too.
        plot.caption.position =  "plot")



case_study_cols <- c(Spectral[c(1, 7, 11)], 
                     adjustcolor(Greys[2], .25))

names(case_study_cols) <- c("Na h-Eileanan Siar", 
                            "Northumberland",
                            "Perth and Kinross", 
                            "Other")
```

# Data
Data on deprivation comes from the English Indix of Multiple Deprivation (https://www.gov.uk/government/statistics/english-indices-of-deprivation-2019) and the Scottish Index of Multiple Deprivation (https://simd.scot/2016/#/simd2016/BTTTFTT/9/-4.0000/55.9000/). These data are aggregated at a similar small-area output (SMO), which allows us to get more granularity from the data across England and Scotland. 

The scores, which are used in calculating the index, are used -- ***not the index ranking***. For instance, one score used to determine income deprivation is the percentatge of people within a SMO that are income deprived. Index rankings themselves are not appropriate for comparing places across the two indexes as they are relational within countries (not between). 


## Employment Deprivation 
```{r employment-overview-scratch, echo=F, warning=F, message=F}


dep_england_tidy <- england_dep_df %>% 
  select(name = local_authority_district_name_2019, 
         income_rate = income_score_rate, 
         employment_rate = employment_score_rate) %>% 
  mutate(case_study = case_when(
    str_detect(name, 'Northumberland')~'Northumberland', 
    T ~ 'Rest of England-Scotland'
  )) 



dep_scotland_tidy <- scotland_dep_df %>% 
  select(name = council_area, 
         employment_rate, 
         income_rate) %>% 
  mutate_if(is.factor, as.character) %>% 
  mutate(employment_rate = parse_number(employment_rate)/100, 
         income_rate = parse_number(income_rate)/100) %>% 
  mutate(case_study = case_when(
    str_detect(name, 'Perth_and_Kinross')~'Perth and Kinross', 
    str_detect(name, 'Na_h-Eileanan_an_Iar')~'Na h-Eileanan Siar',
    T ~ 'Rest of England-Scotland'
  
  )) 


income_employ_dep <- 
  dep_england_tidy %>% 
  bind_rows(dep_scotland_tidy)

```

```{r employment-map-scratch, echo=FALSE, warning=F, message=F}



northumberland_sf <-  dep_england %>% 
    filter(str_detect(lso11nm, 'Northumberland'))


scotland_sf <- dep_scotland %>% 
    select_if(is.factor) %>% 
    mutate_at(vars(incm_rt:ncntrlht_r), 
              list(~parse_number(as.character(.))/100))


perthshire <- scotland_sf %>% 
    filter(str_detect(concl_r, 
                      'Perth_and_Kinross'))  %>% 
    select(Name, contains('emp'), 
           income_rate = incm_rt, 
           over_crowd_rate = ovrcrwdd_r) %>% 
    rename(employment_rate = emplymnt_r)


Na_h_Eileanan_an_Iar <- scotland_sf %>% 
    filter(str_detect(concl_r, 
                      'Na_h-Eileanan_an_Iar'))  %>% 
    select(Name, contains('emp'), 
           income_rate = incm_rt, 
           over_crowd_rate = ovrcrwdd_r) %>% 
    rename(employment_rate = emplymnt_r)

Na_h_Eileanan_an_Iar_simp <- rmapshaper::ms_simplify(Na_h_Eileanan_an_Iar)


northumberland_employment <- northumberland_sf %>% 
    select(Name = lso11nm, contains('emp'), 
           income_rate = incm_dprvt) %>% 
    rename(employment_rate = emp__)


<<<<<<< HEAD
union_ls <- map(list(perthshire, 
                     Na_h_Eileanan_an_Iar,    north_tyneside_employment), 
                function(x){
   st_union(x)
}) %>% 
    set_names(c('perthshire_union', 
                'nah_eileanan_union', 
                'north_tyneside_union'))

=======
>>>>>>> 15fdc92a8343fff8c1805198b3b61445fcc1591e

place_names <- c('Perth and Kinross', 
       'Outer Hebrides', 
       'Northumberland')


```

### Employment deprivation definitions
<<<<<<< HEAD

***English Indeces of Multiple Deprivation*** - The Employment Deprivation Domain measures the proportion of the working-age population in an area involuntarily excluded from the labour market. This includes people who would like to work but are unable to do so due to unemployment, sickness or disability, or caring responsibilities.
=======
>>>>>>> 15fdc92a8343fff8c1805198b3b61445fcc1591e


```{r employment-table, echo=F, warning=F, message=F}
income_employ_dep %>% 
  group_by(case_study) %>% 
  summarise(Median = median(employment_rate, na.rm = T), 
          Mean = mean(employment_rate, na.rm = T), 
          Minimum = min(employment_rate, na.rm = T), 
          Maximum = max(employment_rate, na.rm = T)) %>% 
  mutate_if(is.numeric, list(~percent(.))) %>% 
  rename('Case study area' = case_study) %>% 
  flextable(cwidth = c(2,1,1,1,1)) %>% 
  set_caption('Proportion of Employment Deprivation in case study areas')
```


<<<<<<< HEAD
```{r employment-map, echo=FALSE, warning=F, message=F, fig.width=10, fig.height=10}
demployment_ls <- list(perthshire, 
                      Na_h_Eileanan_an_Iar,
                      north_tyneside_employment)
=======
```{r employment-map, echo=FALSE, warning=F, message=F, fig.width=16, fig.height=6}

employment_ls <- list(perthshire,
                      Na_h_Eileanan_an_Iar,
                      northumberland_employment)
>>>>>>> 15fdc92a8343fff8c1805198b3b61445fcc1591e

tmap_style("gray")
title_ls <- glue('{place_names}')


<<<<<<< HEAD
perth_unemployment <- tm_shape(employment_ls[[1]]) +
    tm_fill('employment_rate',
                palette = 'YlGnBu', 
            title  = 'Unemployment Rate', 
            legend.format=
        list(fun=function(x){paste0(x*100, "%")})) +
    tm_shape(union_ls[[1]])+
    tm_borders()+
    tm_scale_bar()+
    tm_layout(title = title_ls[[1]], 
              legend.frame = T)
    
eileanan_unemployment <- tm_shape(employment_ls[[2]]) +
    tm_fill('employment_rate',
                palette = 'YlGnBu', 
            title  = 'Unemployment Rate', 
            legend.format=
              list(fun=function(x) {paste0(x*100, "%")})) +
    tm_shape(union_ls[[2]])+
    tm_borders()+
    tm_scale_bar()+
    tm_layout(title = title_ls[[2]], legend.frame = T)

north_tyneside_employment <- tm_shape(employment_ls[[3]]) +
    tm_fill('employment_rate',
                palette = 'YlGnBu', 
            title  = 'Unemployment Rate', 
            legend.format=
        list(fun=function(x) {paste0(x*100, "%")})) +
    tm_shape(union_ls[[3]])+
    tm_borders()+
    tm_scale_bar()+
    tm_layout(title = title_ls[[3]], legend.frame = T)
         
tmap_arrange(perth_unemployment, 
             eileanan_unemployment, 
             north_tyneside_employment)
=======
employment_maps <- map(1:3, function(x){
  tm_shape(employment_ls[[x]]) +
    tm_fill('employment_rate',
            title  = 'Unemployment Rate',
            palette = '-YlGnBu',
            style = 'log10',
            legend.format=
                list(fun=function(x) {paste0(round(x, 2)*100, "%")})) +
    tm_scale_bar()+
    tm_layout(title = title_ls[[x]],
              legend.frame = T,
              legend.bg.color = 'white',
              asp = 1)
})

tmap_arrange(employment_maps)


employment_map_path <- glue('/Users/emeador/OneDrive - SRUC/Twenty_five_years_on/png/maps/employment/{place_names}_income.png')
>>>>>>> 15fdc92a8343fff8c1805198b3b61445fcc1591e

map2(employment_maps, employment_map_path, function(x, y){
  tmap_save(x,
            filename = y, 
              width = 4.5, 
            height = 4.5)
}) 

``` 


## Income Deprivation 

```{r income-table, echo=F, warning=F, message=F}
dep_england_tidy %>% 
  bind_rows(dep_scotland_tidy) %>% 
  group_by(case_study) %>% 
  summarise(Median = 
              median(income_rate, 
                     na.rm = T), 
            Mean = 
              mean(income_rate, 
                        na.rm = T), 
            Minimum = 
              min(income_rate, 
                  na.rm = T), 
            Maximum = 
              max(income_rate, 
                  na.rm = T)) %>% 
  mutate_if(is.numeric, list(~percent(.))) %>% 
  rename('Case study area' = case_study) %>% 
  flextable(cwidth = c(2,1,1,1,1)) %>% 
  set_caption('Proportion of Employment Deprivation in case study areas')


```


<<<<<<< HEAD
```{r income-map, echo=FALSE, warning=F, message=F, fig.width=10, fig.height=10}
title_ls <- glue('{place_names} - Income Deprivation Rate')

    
perth_income <- 
  tm_shape(employment_ls[[1]]) +
    tm_fill('income_rate',
                palette = 'YlGnBu', 
                style=c("fixed"), 
                breaks = seq(0, .6, .05), 
            title  = 'Income Rate', 
            legend.format=
                list(fun=function(x) {paste0(x*100, "%")})) +
    tm_shape(union_ls[[1]])+
    tm_borders()+
    tm_scale_bar()+
    tm_layout(title = title_ls[[1]], legend.frame = T)
    
eileanan_income <- tm_shape(employment_ls[[2]]) +
    tm_fill('income_rate',
                palette = 'YlGnBu', 
                style=c("fixed"), 
                breaks = seq(0, .6, .05), 
            title  = 'Income Rate', 
            legend.format=
                list(fun=function(x) {paste0(x*100, "%")})) +
    tm_shape(union_ls[[2]])+
    tm_borders()+
    tm_scale_bar()+
    tm_layout(title = title_ls[[2]], legend.frame = T)
=======
```{r income-map, echo=FALSE, warning=F, message=F, fig.width=16, fig.height=6}

title_ls <- glue('{place_names}')

>>>>>>> 15fdc92a8343fff8c1805198b3b61445fcc1591e


employment_ls <- list(perthshire,
                      Na_h_Eileanan_an_Iar,
                      northumberland_employment)
tmap_style("gray")

income_maps <- map(1:3, function(x){
  tm_shape(employment_ls[[x]]) +
    tm_fill('income_rate',
            title  = 'Income Deprivation\nRate',
            palette = '-YlGnBu',
            style = 'log10',
            legend.format=
                list(fun=function(x) {paste0(round(x, 2)*100, "%")})) +
    tm_scale_bar()+
    tm_layout(title = title_ls[[x]],
              legend.frame = T,
              legend.bg.color = 'white',
              asp = 1)
})

tmap_arrange(income_maps)

 income_map_path <- glue('/Users/emeador/OneDrive - SRUC/Twenty_five_years_on/png/maps/income/{place_names}_income.png')

 map2(income_maps, income_map_path, function(x, y){
   tmap_save(x,
             filename = y,
             width = 4.5,
             height = 4.5)
 })

```


# Correlations between poverty and other socioeconomic factors
```{r scotland-correlations, echo=F, message=F, warning=F}

scotland_dep_df %>% 
  select(data_zone, 
         council_area, 
         contains('rate')) %>% 
  mutate_at(vars(contains('rate')),~ parse_number)


```










