---
title: "Deprivation"
author: "Elliot Meador"
date: "22/10/2019"
output: word_document
editor_options: 
  chunk_output_type: console
---

```{r setup, echo=FALSE, warning=F, message=F}
knitr::opts_chunk$set(echo = TRUE)
```
```{r data-library, echo=F, message=F, warning=F, results='hide'}
library(tidyverse)
library(janitor)
library(openxlsx)
library(glue)
library(flextable)
library(scales)
library(ggrepel)

library(sf)
library(tmap)
library(tmaptools)

if (Sys.info()[[1]] == 'Windows') {
    ## Deprivation Data

    dep_england <- st_read('C:/Users/emeador/OneDrive - SRUC/Data/deprivation/england/english_deprivation_score_sf.shp')
    
    dep_scotland <- st_read('C:/Users/emeador/OneDrive - SRUC/Data/deprivation/scotland/scottish_deprivation_score_sf.shp')
    

    source('C:/Users/emeador/OneDrive - SRUC/all_functions.R')
    
} else {

        dep_england <- st_read('/Users/johne.meador/OneDrive - SRUC/Data/deprivation/england/english_deprivation_score_sf.shp')
    
    dep_scotland <- st_read('/Users/johne.meador/OneDrive - SRUC/Data/deprivation/scotland/scottish_deprivation_score_sf.shp')

    
    
    source('/Users/johne.meador/OneDrive - SRUC/all_functions.R')
    
}



```


# Data
Data on deprivation comes from the English Indix of Multiple Deprivation (https://www.gov.uk/government/statistics/english-indices-of-deprivation-2019) and the Scottish Index of Multiple Deprivation (https://simd.scot/2016/#/simd2016/BTTTFTT/9/-4.0000/55.9000/). These data are aggregated at a similar small-area output (SMO), which allows us to get more granularity from the data across England and Scotland. 

The scores, which are used in calculating the index, are used -- ***not the index ranking***. For instance, one score used to determine income deprivation is the percentatge of people within a SMO that are income deprived. Index rankings themselves are not appropriate for comparing places across the two indexes as they are relational within countries (not between). 

We cover employment and income deprivation in the following sections. 
```{r employment-overview-scratch, echo=F, warning=F, message=F}

scotland_dep_df <- read_csv('/Users/johne.meador/OneDrive - SRUC/Data/deprivation/scotland/simd2016_withgeog/simd2016_withinds.csv') %>% 
  clean_names()

dep_scotland_tidy <- scotland_dep_df %>% 
  select(name = council_area, 
         employment_rate, 
         income_rate) %>% 
  mutate_if(is.factor, as.character) %>% 
  mutate(employment_rate = parse_number(employment_rate)/100, 
         income_rate = parse_number(income_rate)/100) %>% 
  mutate(case_study = case_when(
    str_detect(name, 'Perth')~'Perth and Kinross', 
    str_detect(name, 'Na_h-Eileanan_an_Iar')~'Outer Hebrides',
    T ~ 'Other'
  
  )) %>% 
  arrange(case_study)

dep_scotland_tidy %>% 
  count(name, sort = T)


dep_scotland_tidy %>% 
  ggplot(aes(employment_rate, income_rate, color = case_study))+
  geom_point()



```


```{r employment-map-scratch, echo=FALSE, warning=F, message=F}

north_tyneside_sf <-  dep_england %>% 
    filter(str_detect(lso11nm, 'North Tyneside'))


scotland_sf <- dep_scotland %>% 
    select_if(is.factor) %>% 
    mutate_at(vars(incm_rt:ncntrlht_r), 
              list(~parse_number(as.character(.))/100))


perthshire<- scotland_sf %>% 
    filter(str_detect(concl_r, 
                      'Perth_and_Kinross'))  %>% 
    select(Name, contains('emp'), 
           income_rate = incm_rt, 
           over_crowd_rate = ovrcrwdd_r) %>% 
    rename(employment_rate = emplymnt_r)


Na_h_Eileanan_an_Iar<- scotland_sf %>% 
    filter(str_detect(concl_r, 
                      'Na_h-Eileanan_an_Iar'))  %>% 
    select(Name, contains('emp'), 
           income_rate = incm_rt, 
           over_crowd_rate = ovrcrwdd_r) %>% 
    rename(employment_rate = emplymnt_r)


north_tyneside_employment <- north_tyneside_sf %>% 
    select(Name = lso11nm, contains('emp'), 
           income_rate = incm_dprvt) %>% 
    rename(employment_rate = emp__)


union_ls <- map(list(perthshire, 
                     Na_h_Eileanan_an_Iar,    north_tyneside_employment), 
                function(x){
    st_union(x)
}) %>% 
    set_names(c('perthshire_union', 
                'nah_eileanan_union', 
                'north_tyneside_union'))


place_names <- c('Perthsire', 
       'Nah Eileanan an Iar', 
       'North Tyneside')


```




```{r employment-map, echo=FALSE, warning=F, message=F, fig.width=10, fig.height=10}

employment_ls <- list(perthshire, Na_h_Eileanan_an_Iar, north_tyneside_employment)


title_ls <- glue('{place_names} - Employment Rate')

    
perth_unemployment <- tm_shape(employment_ls[[1]]) +
    tm_fill('employment_rate',
                palette = 'YlGnBu', 
                style=c("fixed"), 
                breaks = seq(0, .6, .05), 
            title  = 'Unemployment Rate', 
            legend.format=
                list(fun=function(x) {paste0(x*100, "%")})) +
    tm_shape(union_ls[[1]])+
    tm_borders()+
    tm_scale_bar()+
    tm_layout(title = title_ls[[1]], legend.frame = T)
    
eileanan_unemployment <- tm_shape(employment_ls[[2]]) +
    tm_fill('employment_rate',
                palette = 'YlGnBu', 
                style=c("fixed"), 
                breaks = seq(0, .6, .05), 
            title  = 'Unemployment Rate', 
            legend.format=
                list(fun=function(x) {paste0(x*100, "%")})) +
    tm_shape(union_ls[[2]])+
    tm_borders()+
    tm_scale_bar()+
    tm_layout(title = title_ls[[2]], legend.frame = T)



north_tyneside_employment <- tm_shape(employment_ls[[3]]) +
    tm_fill('employment_rate',
                palette = 'YlGnBu', 
                style=c("fixed"), 
                breaks = seq(0, .6, .05), 
            title  = 'Unemployment Rate', 
            legend.format=
                list(fun=function(x) {paste0(x*100, "%")})) +
    tm_shape(union_ls[[3]])+
    tm_borders()+
    tm_scale_bar()+
    tm_layout(title = title_ls[[3]], legend.frame = T)
         
 tmap_arrange(perth_unemployment, 
             eileanan_unemployment, 
             north_tyneside_employment)

```




## Income Deprivation 
```{r income-map, echo=FALSE, warning=F, message=F, fig.width=10, fig.height=10}

title_ls <- glue('{place_names} - Income Deprivation Rate')

    
perth_income <- tm_shape(employment_ls[[1]]) +
    tm_fill('income_rate',
                palette = 'YlGnBu', 
                style=c("fixed"), 
                breaks = seq(0, .6, .05), 
            title  = 'Income Rate', 
            legend.format=
                list(fun=function(x) {paste0(x*100, "%")})) +
    tm_shape(union_ls[[1]])+
    tm_borders()+
    tm_scale_bar()+
    tm_layout(title = title_ls[[1]], legend.frame = T)
    
eileanan_income <- tm_shape(employment_ls[[2]]) +
    tm_fill('income_rate',
                palette = 'YlGnBu', 
                style=c("fixed"), 
                breaks = seq(0, .6, .05), 
            title  = 'Income Rate', 
            legend.format=
                list(fun=function(x) {paste0(x*100, "%")})) +
    tm_shape(union_ls[[2]])+
    tm_borders()+
    tm_scale_bar()+
    tm_layout(title = title_ls[[2]], legend.frame = T)



north_tyneside_income <- tm_shape(employment_ls[[3]]) +
    tm_fill('income_rate',
                palette = 'YlGnBu', 
                style=c("fixed"), 
                breaks = seq(0, .6, .05), 
            title  = 'Income Rate', 
            legend.format=
                list(fun=function(x) {paste0(x*100, "%")})) +
    tm_shape(union_ls[[3]])+
    tm_borders()+
    tm_scale_bar()+
    tm_layout(title = title_ls[[3]], legend.frame = T)
         
 
tmap_arrange(perth_income, 
             eileanan_income, 
             north_tyneside_income)
```











